Constructing a universe for the setoid model

### 2020-01-08 types20-abstract.tex
Fix some macros

We can model this as an inductive type $\ad{in-U}$ that \emph{carves out} all
types in $\Set$ that are in the image of $\El$:
%
\begin{align*}
  & \textsf{data}\ \ad{in-U} : \Set \to \Set_1 \\
  & \quad \textsf{inBool} : \textsf{in-U}\ \mathbbm{2} \\
  & \quad \textsf{inPi}
  :  \{A : \Set\} \{B : A \to \Set\}\\
  & \ \qquad \to \textsf{in-U}\ A
  \to ((a : A) \to \textsf{in-U}\ (B\ a))
  \to \textsf{in-U}\ ((a : A) \to (B\ a))
\end{align*}
\begin{align*}
  \U & = \Sigma (A : \Set) (\textsf{in-U}\ A) \\
  \El & = \textsf{proj}_1
\end{align*}

Note that this construction gives rise to an inductive-recursive universe in
$\Set_1$, rather than $\Set$, thus the meta-theory must contain at least one
universe. Our result is that a modified form of this translation also works for
the more complex inductive-recursive type we need to model the universe of
setoids. 
%
In particular, in addition to $\ad{in-U}$ for defining $\U$ as before, we also
introduce a family $\ad{in-U∼}$ of binary relations between types in the
universe, from which we then define $\_\sim_{\setoidU}\_$.
%
However, we now need an inductive-inductive type in the target theory.
%
\begin{align*}
  & \textsf{data}\ \ad{in-U} : \Set \to \Set_1 \\
  & \textsf{data}\ \ad{in-U∼} : \{A\ A' : \Set\} \to \ad{in-U}\ A \to \ad{in-U}\ A' \to (A \to A' \to \Prop) \to \Set_1 \\
  & \setoidU : \Set_1 \\
  & \El : \setoidU \to \Set \\
  & \_\sim_{\setoidU}\_ : \setoidU \to \setoidU \to \Prop_1 \\
  & \_\vdash\_\sim_{\El}\_ : \{a\ a' : \U\} \to a\ \sim_{\setoidU}\ a' \to \El\ a \to \El\ a' \to \Prop \\
  & \\
  & \setoidU = \Sigma\ (X : \Set)\ (\AgdaDatatype{in-U}\ X) \\
  & \El = \textsf{proj}_1 \\
  & (X , p) \sim_{\setoidU} (X' , p') =
   \ \parallel \Sigma\ (R : X \to X' \to \Prop) (\ad{in-U∼}\ p\ p'\ R) \parallel
\end{align*}

where $\parallel A \parallel\ : \Prop$ is propositional truncation for any $A :
\Set$. Intuition would suggest to define $\_\vdash\_\sim_{\El}\_$ by projecting
the relation out of the proof of $a \sim_{\setoidU} a'$, in much the same way as
$\El$ is defined by projecting out of $\U$. That is, $p \vdash x \sim_{\El} y =
\ad{proj}_1\ p\ x\ y$.
%
However, this doesn't work since the type of $p$ is propositionally truncated,
hence it cannot be used to construct a proof-relevant object. We can work around
this by defining $\_\vdash\_\sim_{\El}\_$ explicitly for all type formers, by
induction on the codes $a\ a' : \U$, mutually with a proof that this definition
is logically equivalent to what we would obtain if we could project out of the
truncation.

>>>>

We can model this as an inductive type $\ad{in-U}$ that \emph{carves out} all
types in $\Set$ that are in the image of $\El$:
%
\begin{align*}
  & \textsf{data}\ \ad{in-U} : \Set \to \Set_1 \\
  & \quad \textsf{inBool} : \textsf{in-U}\ \mathbbm{2} \\
  & \quad \textsf{inPi}
  :  \{A : \Set\} \{B : A \to \Set\}\\
  & \ \qquad \to \ad{in-U}\ A
  \to ((a : A) \to \ad{in-U}\ (B\ a))
  \to \ad{in-U}\ ((a : A) \to (B\ a))
\end{align*}
\begin{align*}
  \U & = \Sigma (A : \Set) (\textsf{in-U}\ A) \\
  \El & = \textsf{proj}_1
\end{align*}

Note that this construction gives rise to an inductive-recursive universe in
$\Set_1$, rather than $\Set$, thus the meta-theory must contain at least one
universe. Our result is that a modified form of this translation also works for
the more complex inductive-recursive type we need to model the universe of
setoids. 
%
In particular, in addition to $\ad{in-U}$ for defining $\U$ as before, we also
introduce a family $\ad{in-U∼}$ of binary relations between types in the
universe, from which we then define $\_\sim_{\setoidU}\_$.
%
However, we now need an inductive-inductive type in the target theory.
%
\begin{align*}
  & \textsf{data}\ \ad{in-U} : \Set \to \Set_1 \\
  & \textsf{data}\ \ad{in-U∼} : \{A\ A' : \Set\} \to \ad{in-U}\ A \to \ad{in-U}\ A' \to (A \to A' \to \Prop) \to \Set_1 \\
  & \setoidU : \Set_1 \\
  & \El : \setoidU \to \Set \\
  & \_\sim_{\setoidU}\_ : \setoidU \to \setoidU \to \Prop_1 \\
  & \_\vdash\_\sim_{\El}\_ : \{a\ a' : \U\} \to a\ \sim_{\setoidU}\ a' \to \El\ a \to \El\ a' \to \Prop \\
  & \\
  & \setoidU = \Sigma\ (X : \Set)\ (\ad{in-U}\ X) \\
  & \El = \textsf{proj}_1 \\
  & (X , p) \sim_{\setoidU} (X' , p') =
   \ \parallel \Sigma\ (R : X \to X' \to \Prop) (\ad{in-U∼}\ p\ p'\ R) \parallel
\end{align*}

where $\parallel A \parallel\ : \Prop$ is propositional truncation for any $A :
\Set$. Intuition would suggest to define $\_\vdash\_\sim_{\El}\_$ by projecting
the relation out of the proof of $a \sim_{\setoidU} a'$, in much the same way as
$\El$ is defined by projecting out of $\U$. That is, $p \vdash x \sim_{\El} y =
\ad{proj}_1\ p\ x\ y$.
%
However, this doesn't work since the type of $p$ is propositionally truncated,
hence it cannot be used to construct a proof-relevant object. We can work around
this by defining $\_\vdash\_\sim_{\El}\_$ explicitly for all type formers, by
induction on the codes $a\ a' : \setoidU$, mutually with a proof that this
definition is logically equivalent to what we would obtain if we could project
out of the truncation.
